---
title: "time series modeling"
output: pdf_document
urlcolor: blue
geometry: margin=0.75in
---

Install and load required packages
```{r, message=FALSE}
if (!require("quantmod")) install.packages("quantmod") 
if (!require("xts")) install.packages("xts") 
if (!require("ggplot2")) install.packages("ggplot2") 
if (!require("ggpubr")) install.packages("ggpubr") 
if (!require("tseries")) install.packages("tseries") 
if (!require("urca")) install.packages("urca") # Johansen cointegration test
if (!require("tsDyn")) install.packages("tsDyn") 
if (!require("timetk")) install.packages("timetk")
if (!require("dplyr")) install.packages("dplyr")
library(quantmod); library(xts); library(ggplot2); library(ggpubr); library(tseries) 
library(urca); library(tsDyn); library(timetk); library(dplyr)
```

# 1. Data

Canadian and US macroeconomic data: GDP growth, unemployment, interest rate, inflation rate
```{r, warning=FALSE}
table <- read.csv(file="Macro_data_can.csv", header=TRUE, sep=",")
ind <- as.Date(table$Index, format="%Y-%m-%d")
table <- subset(table, select=-c(Index))
Macro_data_can <- xts(x=table, order.by=ind)

Macro_data <- Macro_data_can
Macro_data <- na.omit(Macro_data) # balance the panel
Macro_data <- window(Macro_data, start="2012-01-01") # starting date for data
vnames <- names(Macro_data) 
dat <- as.matrix(na.omit(diff(Macro_data)))
```

# 2. Model Selection

VEC model: Johansen cointegration test for combined Canadian and US data
```{r}
cointegration <- ca.jo(Macro_data, type="trace", ecdet="trend", spec="transitory")
sc <- summary(cointegration)
sc
```

# 3.m  VAR, TVAR, VEC time-series validation MSE 
```{r, warning=FALSE}
for (lag in 1:3) {
  print("lag:")
  print(lag)

  for (m in 1:3) {
    
    TT <- nrow(dat)
    T1 <- floor(0.5*TT) # start at 50% of the sample size
    step <- 12 # forecast data horizon for MSE
    tseq <- seq(from=T1, to=TT, by=step)
    tseq <- tseq[-length(tseq)]
    MSE.t <- matrix(0,nrow=tseq[length(tseq)]+step-T1,ncol=length(vnames)) # initialize
    colnames(MSE.t) <- vnames
  
    for (j in tseq) {
  
      # VAR model
      if (m==1) {model <- lineVar(data=dat[1:j-1,], lag=lag, model="VAR", I="diff")
                 fcst <- predict(model, n.ahead=step)}
  
      # TVAR model
      if (m==2) {model <- TVAR(data=dat[1:j-1,], lag=lag, model="TAR", nthresh=1, trace=F)
                 fcst <- predict(model, n.ahead=step)}    
      
      # VEC model
      if (m==3) {model <- lineVar(data=dat[1:j-1,], lag=lag, r=1, model="VEC")
                 fcst <- predict(model, n.ahead=step)}    
      #Note: TVEC model is not implemented in R for more than 2 variables
  
      js <- j+step-1
      MSE.t[(j-T1+1):(js-T1+1),] <- (dat[j:js,]-fcst)^2
    }
  
    if (m==1) print("VAR")
    if (m==2) print("TVAR")  
    if (m==3) print("VEC")
  
    MSE <- matrix(colMeans(MSE.t), nrow=1)
    colnames(MSE) <- vnames
    print(MSE)
    print(" ")
  }
}
```
# 4. Smallest Time Series Validation MSE
TVAR yields the smallest time series validation MSE for all variables at lag 1."
TVAR yields the smallest time series validation MSE for all variables except CPI at lag 2 and 3.
VAR yields the smallest time series validation MSE for CPI at lag 2 and 3.

